---
title: "p8105_hw1_sz3030"
author: "Shaocong Zhang"
date: "9/27/2021"
output: github_document
---

# Problem 1

```{r problem_1}
library(tidyverse)

prob1_tab = tibble(
  norm_samp = rnorm(10),
  log_vec = norm_samp > 0,
  char_vec = c("a", "b", "c", "d", "e", "f", "g", 'h', "i", "j"),
  fac_vec = factor(c("easy", "median", "hard", "easy", "median", "hard", "easy", "median", "hard", "easy"))
)

# mean of the random sample
mean(pull(prob1_tab, norm_samp))
# mean of the logical vector
mean(pull(prob1_tab, log_vec))
# mean of the character vector
mean(pull(prob1_tab, char_vec))
# mean of the factor vector
mean(pull(prob1_tab, fac_vec))
```

Based the mean values of each variable, we can get that the character vector and the factor vector don't work while the random sample and logical vector work.

```{r eval= FALSE}
# use as.numeric function to the logical variable
as.numeric(pull(prob1_tab, log_vec))
# use as.numeric function to the character variable
as.numeric(pull(prob1_tab, char_vec))
# use as.numeric function to the factor variable
as.numeric(pull(prob1_tab, fac_vec))
```

The logical and factor variables are successfully converted to the numeric variables while the character variable still cannot be converted.

The logical variable is thought to be 0 or 1 as default to represent FALSE or TRUE, respectively. So it is not surprising that logical variable has mean value. `mean()` function does work for logical variables.

The factor variable is also arranged specific values to represent different levels in the variable once we use `as.numeric` vector because of the factor variable is ordinal. But there doesn't have a default value in the factor variable so that we cannot get its mean directly.

For the character variable, we cannot use `as.numeric` vector to assign numbers to its characters and therefore it helps explain why we cannot get the mean in the previous chunck.

# Problem 2

```{r }
data("penguins", package = "palmerpenguins")
```

The dataset contains several variables of the penguins from `r min(pull(penguins, year))` to `r max(pull(penguins, year))`, they are `r names(penguins)`. The average body mass of the penguins is `r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), 2)`. There are three species `r levels(pull(penguins, species))` in this dataset and the number of data we collected from the three species are `r summary(pull(penguins, species))`, respectively. What's more, this is a `r nrow(penguins)` X `r ncol(penguins)` dataset and the mean flipper length we get from the dataset is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)`.

```{r warning = FALSE}
# We first pull the X-Y and species variables from the dataset and make a data frame
plot_prob2 = tibble(
  y = penguins$flipper_length_mm,
  x = penguins$bill_length_mm,
  species = penguins$species
)

ggplot(plot_prob2, aes(x = x, y = y, color = species)) + geom_point()
```


```{r warning = FALSE}
ggsave("hw1_scatter_plot.pdf")
```

