---
title: "p8105_hw1_sz3030"
author: "Shaocong Zhang"
date: "9/27/2021"
output: github_document
---

# Problem 1

```{r problem_1}
library(tidyverse)

prob1_tab = tibble(
  norm_samp = rnorm(10),
  log_vec = norm_samp > 0,
  char_vec = c("a", "b", "c", "d", "e", "f", "g", 'h', "i", "j"),
  fac_vec = factor(c("easy", "median", "hard", "easy", "median", "hard", "easy", "median", "hard", "easy"))
)

# mean of the random sample
mean(pull(prob1_tab, norm_samp))
# mean of the logical vector
mean(pull(prob1_tab, log_vec))
# mean of the character vector
mean(pull(prob1_tab, char_vec))
# mean of the factor vector
mean(pull(prob1_tab, fac_vec))
```

Based the mean values of each variable, we can get that the character vector and the factor vector don't work while the random sample and logical vector work.

```{r eval= FALSE}
# use as.numeric function to the logical variable
as.numeric(pull(prob1_tab, log_vec))
# use as.numeric function to the character variable
as.numeric(pull(prob1_tab, char_vec))
# use as.numeric function to the factor variable
as.numeric(pull(prob1_tab, fac_vec))
```

The logical and factor variables are successfully converted to the numeric variables while the character variable still cannot be converted.

The logical variable is thought to be 0 or 1 as default to represent FALSE or TRUE, respectively. So it is not surprising that logical variable has mean value.

The factor variable is also arranged specific values to represent different levels in the variable once we use `as.numeric` vector. But there doesn't have a default value in the factor variable so that we cannot get its mean directly.

For the character variable, we cannot use `as.numeric` vector to assign numbers to its characters and therefore it helps explain why we cannot get the mean in the previous chunck.

